Supplementary Information

### 2020-07-02 manuscript/pnas_style/manuscript.tex
new sentence

Our analyses reveal that a sizable proportion of allele frequency change in
\vb{these experimental evolution populations} is due to the (likely indirect)
action of selection. Capitalizing on replicated designs, we characterized the
extent to which convergent selection pressures lead to parallel changes in
allele frequencies across replicate populations, and found that a substantial
proportion of the response is shared across short timescales.  These likely
represent substantial under-estimates of the contribution of linked selection
because the studies we have reanalyzed do not sequence the population each
generation, preventing us from including the effects of stronger correlations
between adjacent generations. Furthermore, our estimation methods are
intentionally conservative, for example they exclude the contribution of
selection that does not persist across generations and selection that reverses
sign; thus they can be seen as a lower bound of the effects of selection, which
we have confirmed through forward-in-time simulations. Finally, \vb{through
  simulation results, we show that for a given level of additive genetic
  variance, the strengths of temporal and replicate covariances depend on the
  mode of selection, the details of the populations or selection experiment,
  and the level of linkage disequilibria, yet the level of temporal covariance
  is relatively invariant to the number of loci underlying fitness, as long as
fitness is sufficiently polygenic.}

>>>>

Our analyses reveal that a sizable proportion of allele frequency change in
\vb{these experimental evolution populations} is due to the (likely indirect)
action of selection. Capitalizing on replicated designs, we characterized the
extent to which convergent selection pressures lead to parallel changes in
allele frequencies across replicate populations, and found that a substantial
proportion of the response is shared across short timescales.  These likely
represent substantial under-estimates of the contribution of linked selection
because the studies we have reanalyzed do not sequence the population each
generation, preventing us from including the effects of stronger correlations
between adjacent generations. Furthermore, our estimation methods are
intentionally conservative, for example they exclude the contribution of
selection that does not persist across generations and selection that reverses
sign; thus they can be seen as a lower bound of the effects of selection, which
we have confirmed through forward-in-time simulations.  \vb{Future work could
  refine our estimates by including selection's impact on the variance in
  allele frequency terms (e.g. see equation 26 of \textcite{Buffalo2019-io}),
  and possibly quantifying the covariances missed when sequencing is not done
  each generation; both would lead to less conservative estimates that could
  show a large impact of selection.  Finally, through simulation results, we
  show that for a given level of additive genetic variance, the strengths of
  temporal and replicate covariances depend on the mode of selection, the
  details of the populations or selection experiment, and the level of linkage
disequilibria, yet the level of temporal covariance is relatively invariant to
the number of loci underlying fitness, as long as fitness is sufficiently
polygenic.}

### 2020-07-02 manuscript/pnas_style/supp.tex
new sentence

We first investigated the effects of the selected trait's genetic architecture
on temporal covariances and $G(t)$ by neutrally burning in a population for
$10N$ generations, and selecting on the trait with an exponential fitness
function \vb{(where fitness of a trait value $z$ is $w(z) \propto \exp(x)$)}.
The exponential fitness function corresponds to multiplicative selection across
sites and so serves as the simplest directional selection model of a trait to
understand the effects of genetic architecture on the statistics we have used
in the paper.

@@

Overall, we \vb{again see a} finding in \textcite{Buffalo2019-io}: that the
initial expected temporal covariance conditioned on $V_A$, is invariant to the
number of loci determining the trait's value, $L$ (Supplementary Figure
\ref{suppfig:sim-expfit-covs}). We do find some evidence that the decay in
temporal covariance is faster when the trait has a monogenic basis (see the
third column of Supplementary Figure \ref{suppfig:sim-expfit-covs}); this is
expected \vb{since} the selection coefficients are larger for these monogenic
simulations, leading to faster allele frequency changes and a rapid change in
additive genetic variance.

@@

In our analysis of the \textcite{Barghi2019-qy} data, we describe our statistic
$G(t)$ as a lower bound for two reasons: (1) the population is sequenced every
ten generations, meaning the temporal covariances between adjacent generations
cannot contribute to the numerator of $G(t)$ but contributes to the
denominator, and (2) the estimate of $G(t)$ ignores linked selection's
contribution to the per-generation variance in allele frequency change (this is
the difference between our $G(t)$ and $G'(t)$ estimators, the latter of which
\vb{is defined in \textcite{Buffalo2019-io} and} includes these variance terms.
To verify that $G(t)$ estimated every ten generations is indeed a lower bound,
we used a simulation procedure similar to the exponential fitness function
simulations (described in Supplementary Material Section
\ref{supp:genetic-arch}), and calculated the temporal covariances and $G(t)$
both each generation, and every ten generations. Unlike the simulations
described in \ref{supp:genetic-arch} \vb{where selection began at $10N+5$
generations}, selection \vb{starts at generation $10N$ here}, and used trait
$V_A = 0.01$ and targeted $L = 1000$ sites affecting the trait.

First, comparing $G(t)$ when sampling population frequencies every generation
versus every ten generations, we confirm that the ten-generation block $G(t)$
is a lower bound of the $G(t)$ trajectory when sampling is every generation
(red and blue lines in Supplementary Figure \ref{suppfig:supp-blocks}A).
Furthermore, since we control the population size \vb{and reproductive sampling
scheme} in our simulations at $N = 1000$ diploids, we know the drift effective
population size in the absence of selection. This allows us to estimate
$G'(t)$, which is a measure of $G(t)$ that accounts for the linked selection's
inflation of the variance in allele frequency change between two generations
\vb{($G'(t) = 1 - \nicefrac{t\E(p_0(1-p_0))}{2 \hat{N}\var(p_t - p_0)}$  see
equation 26 of \cite{Buffalo2019-io}} for more detail).  Plugging in the drift
effective population size $N_e = 1000$ into the expression for $G'(t)$ and
using the $\var(p_t - p_0)$ calculated for different $t$'s, we see that $G(t)$
calculated every generation does not account for linked selection's inflation
of $\var(\Delta p_t)$ and underestimates the true impact of linked selection as
expected (dashed gray line in Supplementary Figure \ref{suppfig:supp-blocks}A).

>>>>

We first investigated the effects of the selected trait's genetic architecture
on temporal covariances and $G(t)$ by neutrally burning in a population for
$10N$ generations, and selecting on the trait with an exponential fitness
function \vb{(where fitness of a trait value $z$ is $w(z) \propto \exp(z)$)}.
The exponential fitness function corresponds to multiplicative selection across
sites and so serves as the simplest directional selection model of a trait to
understand the effects of genetic architecture on the statistics we have used
in the paper.

@@

Overall, we \vb{again see a} finding of \textcite{Buffalo2019-io}: that the
initial expected temporal covariance conditioned on $V_A$, is invariant to the
number of loci determining the trait's value, $L$ (Supplementary Figure
\ref{suppfig:sim-expfit-covs}). We do find some evidence that the decay in
temporal covariance is faster when the trait has a monogenic basis (see the
third column of Supplementary Figure \ref{suppfig:sim-expfit-covs}); this is
expected \vb{since} the selection coefficients are larger for these monogenic
simulations, leading to faster allele frequency changes and a rapid change in
additive genetic variance.

@@

In our analysis of the \textcite{Barghi2019-qy} data, we describe our statistic
$G(t)$ as a lower bound for two reasons: (1) the population is sequenced every
ten generations, meaning the temporal covariances between adjacent generations
cannot contribute to the numerator of $G(t)$ but contributes to the
denominator, and (2) the estimate of $G(t)$ ignores linked selection's
contribution to the per-generation variance in allele frequency change. \vb{ In
\textcite{Buffalo2019-io}, we define an alternative estimator
that includes selection's effects on these variance terms,

\begin{equation}
  G'(t) = 1 - \frac{t\E(p_0(1-p_0))}{2 {N_e}\var(p_t - p_0)},
\end{equation}
%
which can be calculated when the drift-effective population size $N_e$ can
be estimated (see equation 26 in \cite{Buffalo2019-io} for details).}

To verify that $G(t)$ estimated every ten generations is indeed a lower bound,
we used a simulation procedure similar to the exponential fitness function
simulations (described in Supplementary Material Section
\ref{supp:genetic-arch}), and calculated the temporal covariances and $G(t)$
both each generation, and every ten generations. Unlike the simulations
described in \ref{supp:genetic-arch} \vb{where selection began at $10N+5$
generations}, selection \vb{starts at generation $10N$ here}, and used trait
$V_A = 0.01$ and targeted $L = 1000$ sites affecting the trait.

First, comparing $G(t)$ when sampling population frequencies every generation
versus every ten generations, we confirm that the ten-generation block $G(t)$
is a lower bound of the $G(t)$ trajectory when sampling is every generation
(red and blue lines in Supplementary Figure \ref{suppfig:supp-blocks}A).
Furthermore, since we control the population size \vb{and reproductive sampling
scheme} in our simulations at $N = 1000$ diploids, we know the drift-effective
population size in the absence of selection, \vb{which allows us to estimate
$G'(t)$}. Plugging in the drift-effective population size $N_e = 1000$ into
the expression for $G'(t)$ and using the $\var(p_t - p_0)$ calculated for
different $t$'s, we see that $G(t)$ calculated every generation does not
account for linked selection's inflation of $\var(\Delta p_t)$ and
underestimates the true impact of linked selection as expected (dashed gray
line in Supplementary Figure \ref{suppfig:supp-blocks}A).
